---
title: "Visualization_2"
author: "Lectured by Jeff Goldsmith"
date: "2022-10-04"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(patchwork)
library(ggridges)
```

```{r read data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>% 
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>% 
  select(name, id, everything())
```

# Labels and Scales 

We can make scatterplots and label the plots with axis labels, plot titles, and captions using the `labs()` function.
We can also adjust the scales, including their positions.
```{r plot 1}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package") +
  scale_x_continuous(
    breaks = c(-10, 0, 15),    # can adjust how your scale looks like.
    labels = c("-10C", "0C", "15C")) +  # label your scale.
  scale_y_continuous(
    trans = "sqrt",       # transform the variable by the function of square root.
    position = "right")   # place the y-scale on the right side of the plot.
```
There are many other ways to play around with the plot. For instance, `scale_y_sqrt()` can be added to a ggplot object to transform the Y scale, and `xlim()` can be used to control the plot limits in the X axis.

# Color Scheme

```{r plot 2}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package") +
  scale_color_hue(
    name = "Location",   # changed the variable "name" to "Location".
    h = c(100, 300))  # Not encourage to do this. Color is very ugly.
```

```{r viridis color}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package") +
  viridis::scale_color_viridis(     # This color pallate is good.
    name = "Location",   # changed the variable "name" to "Location".
    discrete = TRUE)
```

# Themes

```{r create a data}
ggp_weather = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package") +
  viridis::scale_color_viridis(     # This color pallate is good.
    name = "Location",   # changed the variable "name" to "Location".
    discrete = TRUE)
```

```{r legend position}
ggp_weather + 
  theme_bw() + # change how the background grid looks like. Try _classic, _minimal, etc.
  theme(legend.position = "bottom")  # change legend position to bottom.
# Must change grid theme first, then legend position later. Otherwise, theme position would not change.
```

# Setting how my plots will look.

Set these options at the beginning of your document, so you don't have to write the same code over and over again in every single code chunk.
```{r setup at very beginning, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Data in geom()

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(waikiki, aes(x = date, y = tmax)) +
  geom_point() +
  geom_line(data = central_park)
```

# Patchwork

```{r}
tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()

prcp_density =
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5)
```

# Data Manipulation

In order to put something on a plot, a variable has to be a factor vector. let's say `name` is character variable, but when we plot, R will change it to factor variable during the plotting process.
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

```{r}
weather_df %>% 
  mutate(name = fct_reorder(name, tmax)) %>%  # reordering "names" by which has the lowest tmax comes first.
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

```{r}
pulse_df =
  haven::read_sas("Data/public_pulse_data.sas7bdat")  %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)
# another way to do this:
# mutate(
#   visit = fct_relevel(visit, "bl"))
ggplot(pulse_df, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```




