---
title: "Visualization_2"
author: "Lectured by Jeff Goldsmith"
date: "2022-10-04"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(patchwork)
library(ggridges)
```

```{r read data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>% 
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>% 
  select(name, id, everything())
```

# Labels and Scales 

We can make scatterplots and label the plots with axis labels, plot titles, and captions using the `labs()` function.
We can also adjust the scales, including their positions.
```{r plot 1}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package") +
  scale_x_continuous(
    breaks = c(-10, 0, 15),    # can adjust how your scale looks like.
    labels = c("-10C", "0C", "15C")) +  # label your scale.
  scale_y_continuous(
    trans = "sqrt",       # transform the variable by the function of square root.
    position = "right")   # place the y-scale on the right side of the plot.
```
There are many other ways to play around with the plot. For instance, `scale_y_sqrt()` can be added to a ggplot object to transform the Y scale, and `xlim()` can be used to control the plot limits in the X axis.

# Color Scheme

`scale_color_hue()` control the color scale and the name in the plot legend. But creating your own color scheme usually doesn’t go well - can be very ugly.
```{r plot 2}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package") +
  scale_color_hue(
    name = "Location",   # changed the variable "name" to "Location".
    h = c(100, 300))  # Not encourage to do this. Color is very ugly.
```

`viridis` package is better.
```{r viridis color}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package") +
  viridis::scale_color_viridis(     # This color palate is good.
    name = "Location",   # changed the variable "name" to "Location".
    discrete = TRUE)
```
##### We used `discrete = TRUE` because the color aesthetic is mapped to a discrete variable. In other cases (for example, when color mapped to `prcp`) you can omit this argument to get a continuous color gradient. The `viridis::scale_fill_viridis()` function is appropriate for the fill aesthetic used in histograms, density plots, and elsewhere.


# Themes

Themes are used to modify non-data elements of a plot – they don’t change mappings or how data are render, but control things like background color and location of the the legend.

```{r create a dataset first}
ggp_weather = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    x = "Minimum Daily Temp (C)",
    y = "Maximum Daily Temp (C)",
    title = "Scatterplot of daily temp extremes",
    caption = "Data come from the rnoaa package") +
  viridis::scale_color_viridis(     
    name = "Location",   
    discrete = TRUE)
```

* Change background grid: `theme_` "bw", "classic", "minimal" and many more.

* Change legend position: `theme(legend.position = "top/bottom/left") *Right is default*.

```{r legend position}
ggp_weather + 
  theme_bw() + # change how the background grid looks like. Try _classic, _minimal, etc.
  theme(legend.position = "bottom")  # change legend position to bottom.
# Must change grid theme first, then legend position later. Otherwise, theme position would not change.
```
Must change grid theme first, then change other theme features later. Otherwise, theme position would not change.

# Learning Assessment 1

Revisit the plot showing tmax against date for each location. Use labels, scale options, and theme changes to improve the readability of this plot.

```{r LA1}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
  geom_smooth(se = FALSE) +   # se = FALSE means don't display CI around smooth line.
  geom_point(aes(size = prcp), alpha = 0.5) +  
  labs(
    title = "Temperature and precipitation plot",
    x = "Date",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package") +
  theme_minimal() +
  theme(legend.position = "bottom")
```


# Setting options

Set these options at the beginning of your document, so you don't have to write the same code over and over again in every single code chunk.
```{r setup at very beginning, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Data argument in `geom()`

We can split weather_df into separate datasets for Central Park and Waikiki. Then we use one in ggplot() and another in geom_line():
```{r two datasets on one plot}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
```

# Patchwork

We can display several plots on the same page (must load `patchwork` package at the beginning):
```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```


# Data Manipulation

In order to put something on a plot, a variable has to be a factor vector. let's say `name` is character variable, but when we plot, R will change it to factor variable during the plotting process.

```{r original order}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```
As we can see in the plot, "Central Park" goes first, then "Waikiki", lastly "Waterhole"


```{r reorder by other variable}
weather_df %>% 
  mutate(name = fct_reorder(name, tmax)) %>%  # reordering "names" by whichever has the lowest tmax comes first.
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Categorical variables will be ordered alphabetically; factors will follow the specified order level that underlies the variable labels. You can change the order level of a factor variable to your specified preference using `forcats::fct_relevel` or according to the value of another variable using `forcats::fct_reorder`.

This way of reordering need to type out all "names" in the order that you want:
```{r relevel by hand}
weather_df %>%
  mutate(name = fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>%  # Don't write 'forcats::' is fine.
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```


How to create a plot that has common x- and -y-axis shared by three different locations?
```{r shared axis}
weather_df %>%
  select(name, tmax, tmin) %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) +    # facet_grid(name ~ .) will show horizontal graph instead of vertical.
  viridis::scale_fill_viridis(discrete = TRUE)
```

Practice using `pulse dataset`:
```{r pulse}
pulse_df =
  haven::read_sas("Data/public_pulse_data.sas7bdat")  %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)
# another way to do this:
# mutate(
#   visit = fct_relevel(visit, "bl"))
ggplot(pulse_df, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```




