---
title: "Visualization I"
author: "Lectured by Jeff Goldsmith"
date: "2022-09-29"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
```

### Prepare the weather dataset

Download the NOAA weather data and make the dataset clean and readable:

```{r NOAA, message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397","USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"),
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>% 
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY",
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>% 
  select(name, id, everything())
```

### Basic Scatterplot

In the code chunk below, nothing will be shown in the graph because we did not specify what type of `geom_`.
```{r nothing show}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

Now if we add `geom_point()`, then R knows what type of graph we want to plot.
```{r scatterplot correct way}
ggplot(weather_df, aes(x = tmin, y = tmax)) + geom_point()
```

But there is a warning message from the code chunk above because there are missing values in the variables.
We can also make the same scatterplot but in a different way (and drop those missing values).
```{r remove missing values}
weather_df %>% 
  drop_na() %>% 
  ggplot(aes(x = tmin, y = tmax)) + geom_point()
```

We can also save the output of `ggplot()` to an object and modify/print it later.
```{r output}
weather_scatterplot = 
  weather_df %>% 
  drop_na() %>% 
  ggplot(aes(x = tmin, y= tmax))

weather_scatterplot + geom_point()
```

### Advanced Scatterplot

The previous scatterplots only show me a bunch of black dots. What if I want to see the distribution of dots by different locations?
```{r color points by variable, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + geom_point()
```

Now we have a nicely looking graph with legend. What about adding a smooth curve?
```{r smooth line, message = FALSE, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + geom_point(aes(color = name)) + geom_smooth()
```

But the code chunk above only show 1 blue line because we put the aesthetic color = name in geom_point.
If we put the aesthetic color in the ggplot, it will show three lines with different colors.
```{r three lines, message = FALSE, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + geom_point() + geom_smooth()
```

Let's make the data points a bit more transparent, so we can see where the dots overlap easily.
`alpha = ` the lower the value, the more transparent it will be.
`se = ` display confidence interval around smooth (TRUE by default).
```{r more transparent, message = FALSE, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE)
```

It seems difficult to view everything on the same graph because everything is overlapping.

Let's make separate panels; Make the same plot but separate by variables -- use `facet_grid`.
```{r facet, message = FALSE, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
# If do facet_grid(name ~ .) will show horizontal graph instead of vertical.
```

Now I have learned enough about the relationship between `tmin` and `tmax`, I want something that shows the time of year and something about precipitation.
```{r add more variables, message = FALSE, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

I can also make a line graph that only shows the relationship between date and maximum temperature:
```{r line graph, warning = FALSE, message = FALSE}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

When you’re making a scatterplot with lots of data, there’s a limit to how much you can avoid overplotting using alpha levels and transparency. In these cases, `geom_hex()`, `geom_bin2d()`, or `geom_density2d()` can be handy.
```{r if we have tons of rows, warning = FALSE, message = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

### Learning Assessment 1
Write a code chain that starts with `weather_df`; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in `geom_smooth()`).
```{r LA1, message = FALSE, warning = FALSE}
LA1_df = weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax_F = tmax * (9 / 5) + 32,
    tmin_F = tmin * (9 / 5) + 32)
ggplot(LA1_df, aes(x = tmin_F, y = tmax_F)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE)
```

### Learning Assessment 2

```{r LA2}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")

ggplot(weather_df) + geom_point(aes(x = tmin, y = tmax, color = "blue"))
```
In the first attempt, we’re defining the color of the points by hand. In the second attempt, we’re implicitly creating a color variable that has the value `blue` everywhere; `ggplot` is then assigning colors according to this variable using the default color scheme.


### Univariate plots

Histograms, barplots, boxplots, violins, ...

##### Histogram

We can show histograms with separate panels:
```{r histogram}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +   
  geom_histogram() +
  facet_grid(. ~ name)
# If put color = name, the color will only show on the outer line, not the entire columns.
```

Or we can do this: The `position = "dodge"` places the bars for each group side-by-side.
```{r bin width, warning = FALSE}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge", binwidth = 2)
```

##### Density plot

The `adjust` parameter in density plots is similar to the `binwidth` parameter in histograms, and it helps to try a few values. Lastly, adding geom_rug() to a density plot can be a helpful way to show the raw data in addition to the density.
```{r density line, warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name )) +
  geom_density(alpha = .4, adjust = 0.5, color = "blue")
# If use 'color = name' will only show color line; If use 'fill = name' will fill the entire space under line, so do alpha = .4 to make the color more transparent.
```

##### Boxplot
```{r boxplot}
weather_df %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_boxplot()
```

###### Violin plot
```{r violin}
weather_df %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.4)
```

##### Ridge plots

Ridge plots showing plots side-by-side instead of stacking all of them on one layer.
```{r density ridges}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges(scale = 0.85)
# Adjust the scale so bigger plots won't overlap.
```

### Learning Assessment 3
Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable.

```{r LA3, message = FALSE, warning = FALSE}
# Histogram
ggplot(weather_df, aes(x = prcp, fill = name)) + 
  geom_histogram(position = "dodge")

# Density plot
ggplot(weather_df, aes(x = prcp)) +
  geom_density(aes(fill = name), alpha = 0.5)

# Boxplot
ggplot(weather_df, aes(x = name, y = prcp, fill = name)) +
  geom_boxplot()

# Violin plot
ggplot(weather_df, aes(x = name, y = prcp, fill = name)) +
  geom_violin(alpha = 0.5)

# Ridgeplot
ggplot(weather_df, aes(x = prcp, y = name, fill = name)) +
  geom_density_ridges(alpha = 0.5, scale = 0.85)
```

### Saving and embedding plots

First, let's save a plot.
```{r save plot, warning = FALSE, message = FALSE}
weather_scatterplot = weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

weather_scatterplot

ggsave(
  file = "results/weather_scatterplot.pdf", 
  plot = weather_scatterplot,
  width = 10, height = 7)
```

Embedding plots:
The size of the figure created by R, which is controlled using two of the three chunk options `fig.width`, `fig.height`, and `fig.asp`. I prefer a common width and plots that are a little wider than they are tall, so I set options to fig.width = 6 and fig.asp = .6. Second is the **size of the figure** inserted into your document, which is controlled using `out.width` or `out.height`. I like to have a little padding around the sides of my figures, so I set `out.width` = "90%".
```{r embedding}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```


