---
title: "Visualization I"
author: "Lectured by Jeff Goldsmith"
date: "2022-09-29"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
knitr::opts_chunk$set(echo = TRUE)
```

# Prepare the weather dataset

Download the NOAA weather data and make the dataset clean and readable:

```{r NOAA, message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397","USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"),
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>% 
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY",
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>% 
  select(name, id, everything())
```

# Basic Scatterplot

```{r scatterplot nothing show}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

```{r scatterplot correct way}
ggplot(weather_df, aes(x = tmin, y = tmax)) + geom_point()
```

Let's make the same scatterplot, but different

```{r remove missing values}
weather_df %>% 
  drop_na() %>% 
  ggplot(aes(x = tmin, y = tmax)) + geom_point()
```

Let's keep making the same plot but different

```{r}
weather_scatterplot = weather_df %>% 
  drop_na() %>% 
  ggplot(aes(x = tmin, y= tmax))

weather_scatterplot + geom_point()
```

Let's fancy this up a little bit

```{r color points by variable}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + geom_point()

# Can also do this
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + geom_point(aes(color = name)) + geom_smooth()
# But only show 1 blue line because we put the aesthetic color = name in geom_point.

# If put the aesthetic color in the ggplot with x, y will show three lines
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + geom_point() + geom_smooth()

# Why do this???
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE)
```

Maybe make separate panels: Make the same plot but separate by variables `facet_grid`

```{r facet}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
# If do facet_grid(name ~ .) will show horizontal graph instead of vertical.
```

`tmax` vs `tmin` is boring, let's spice it up some.

```{r add more variables}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

## Some quick stuff

```{r if we have tons of rows}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

# Univariate plots

Histograms, barplots, boxplots, violins, ...

##### Histogram
```{r histogram}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +   
  geom_histogram() +
  facet_grid(. ~ name)
# If put color = name, the color will only show on the outer line, not the entire columns.
```

##### Density plot
```{r density line}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name )) +
  geom_density(alpha = .3)
# If use color = name will only show color line; If use fill = name will fill the entire under line, so do alpha = .3 to make the color more transparent.
```

##### Boxplot
```{r boxplot}
weather_df %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_boxplot()
```

###### Violin plot
```{r violin}
weather_df %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()
```

##### Density ridges
```{r density ridges}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
# ridges show density plot by stacking them next to each other instead of overlapping all of them in one stack.
```

# Saving and embedding plots

First, let's save a plot.

```{r save plot}
weather_scatterplot = weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

weather_scatterplot

ggsave(
  file = "results/weather_scatterplot.pdf", 
  plot = weather_scatterplot,
  width = 12, height = 10)
```

```{r, fig.width = 10}
# Show the figure wider.
weather_scatterplot
```


